\vspace{4.0\baselineskip}
\textbf{\hspace{2.5cm} \LARGE \textbf{ACKNOWLEDGMENTS}}

\vspace{15mm}
%\noindent\rule{\linewidth}{2pt}
%It is a great pleasure for me to express my respect and deep sense of gratitude to my M.Tech. supervisor \textbf{Dr. Narottam Chand Kaushal}, Associate Professor, Department of Computer Science \& Engineering, National Institute of Technology Hamirpur, for his wisdom, vision, expertise, guidance, enthusiastic involvement during the course of my M.Tech. at NIT Hamirpur. I am specially indebted to him for teaching me both writing and research skills which have proven to be beneficial for my current research and future career. Without his endless efforts, knowledge and patience this research would not have been possible. It has been a valuable learning experience for me to work under his supervision.\\

It is an incredible delight for me to express my regard and profound feeling of appreciation to my M.Tech. supervisor \textbf{Dr. Narottam Chand Kaushal}, Associate Professor, Department of Computer Science \& Engineering, National Institute of Technology Hamirpur, for his shrewdness, vision, ability, direction, eager association throughout my M.Tech. course at NIT Hamirpur. I am exceptionally obliged to him for showing me both composition and research aptitudes which have demonstrated to be gainful for my momentum research and future vocation. Without his unlimited endeavors, information and guidance this research would not have been conceivable. It has been a significant learning background for me to work under his watch.\\

% \par My gratitude is also extended to all the teachers and staff members of \textbf{Computer Science \& Engineering Department and Computer Centre} for their timely help and cooperation extended throughout the course of investigation.\\

\par My appreciation is additionally stretched out to every one of the teachers and staff individuals from \textbf{Computer Science \& Engineering Department and Computer Centre} for their convenient assistance and collaboration stretched out over the span of research.\\

% \par I am obliged to my parents for their moral support, love, encouragement and blessings to complete this task. Without their dedication and dependability I could not have accomplished this task.\\
\par I am obliged to my family and friends for their ethical help, love, consolation and favors to finish this research. Without their devotion and reliability I couldn't have achieved this undertaking.\\


% \par Finally, I am indebted and grateful to the Almighty for helping me in this endeavor.\
\par At last, I am obligated and appreciative to the Almighty for helping me in this endeavor.

\vspace{1.2cm}
\begin{flushright}
	\textbf{(Atul Agarwal)}
\end{flushright}

